---
title: "Teton Temperature Trends"
author: "Gordon Gianniny"
date: "2023-02-21"
output: html_document
---

```{r, echo=FALSE}
#Loading Packages
library(tidyverse) #For data wrangling
library(lubridate) #For working with dates
library(ggplot2) #Plotting
library(RColorBrewer)
library(patchwork)
library(corrplot)
library(broom)
```


## Overview

This script goes contains different analyses testing for trends in stream temperature in the Teton range. It includes the following analyses: 

  1. Using previously cleaned (by Scott) data - basic lm's testing the effects of year, air temp, snow pack on average summer (July-August) temperature (designed after the "real time warming..." paper from the Alps). 
  



## 1. lm's using previously cleaned data:

### A. Data Import and Cleaning

Read in datafiles

```{r}
ak_basin <- read.csv("Temperature/cleaned_csv/ak_basin_south.csv")
middle_teton <- read.csv("Temperature/cleaned_csv/middle_teton.csv")
n_teton <- read.csv("Temperature/cleaned_csv/n_fork_teton.csv")
paintbrush <- read.csv("Temperature/cleaned_csv/paintbrush.csv")
s_cascade <- read.csv("Temperature/cleaned_csv/s_cascade.csv")
wind_cave <- read.csv("Temperature/cleaned_csv/wind_cave.csv")
```

Re-format data: 

```{r}
#AK Basin:

ak_basin_pivot <- ak_basin %>%
 select(Date, temp_2015:temp_2022)%>% #remove unnecessary cols
  pivot_longer(!c(Date), names_to = "meas_date", values_to = "ak_basin_temp")%>% #pivot longer - make new columns for year and temp, one row for each date for each year
  mutate(year = substr(meas_date,6, 9), #remove "temp_" at beginning of each year to get year column
         dmy_date = paste(Date, year, sep = "-"), #create intermediate date column with year, month, and day
         date_final = as.Date(dmy_date, format = "%d-%b-%Y") #convert intermediate to new date column with as.Date
         )%>% 
  select(-meas_date, -dmy_date, -Date)#%>% #remove intermediate/redundant date columns
  #filter(!is.na(date_final)) #remove extra rows with no date

#Middle Teton

middle_teton_pivot <- middle_teton %>%
  select(Date, temp_2015:temp_2022)%>% #remove unnecessary cols
  pivot_longer(!c(Date), names_to = "meas_date", values_to = "mid_teton_temp")%>% #pivot longer - make new columns for year and temp, one row for each date for each year
  mutate(year = substr(meas_date,6, 9), #remove "temp_" at beginning of each year to get year column
         dmy_date = paste(Date, year, sep = "-"), #create intermediate date column with year, month, and day
         date_final = as.Date(dmy_date, format = "%d-%b-%Y") #convert intermediate to new date column with as.Date
         )%>% 
  select(-meas_date, -dmy_date, -Date)%>% #remove intermediate/redundant date columns
  filter(!is.na(date_final)) #remove extra rows with no date

# N Teton

n_teton_pivot <- n_teton %>%
  select(Date, temp_2015:temp_2022)%>% #remove unnecessary cols
  pivot_longer(!c(Date), names_to = "meas_date", values_to = "n_teton_temp")%>% #pivot longer - make new columns for year and temp, one row for each date for each year
  mutate(year = substr(meas_date,6, 9), 
         dmy_date = paste(Date, year, sep = "-"), 
         date_final = as.Date(dmy_date, format = "%d-%b-%Y")
         )%>% #remove "temp_" at beginning of each year
  select(-meas_date, -dmy_date, -Date)%>%
  filter(!is.na(date_final)) #remove extra rows with no date

# Paintbrush

paintbrush_pivot <- paintbrush %>%
 select(Date, temp_2015:temp_2022)%>% #remove unnecessary cols
  pivot_longer(!c(Date), names_to = "meas_date", values_to = "paintbrush_temp")%>% #pivot longer - make new columns for year and temp, one row for each date for each year
  mutate(year = substr(meas_date,6, 9), 
         dmy_date = paste(Date, year, sep = "-"), 
         date_final = as.Date(dmy_date, format = "%d-%b-%Y")
         )%>% #remove "temp_" at beginning of each year
  select(-meas_date, -dmy_date, -Date)%>%
  filter(!is.na(date_final)) #remove extra rows with no date

# S cascade:

s_cascade_pivot <- s_cascade  %>%
  select(Date, temp_2015:temp_2022)%>% #remove unnecessary cols
  pivot_longer(!c(Date), names_to = "meas_date", values_to = "s_cascade_temp")%>% #pivot longer - make new columns for year and temp, one row for each date for each year
  mutate(year = substr(meas_date,6, 9), 
         dmy_date = paste(Date, year, sep = "-"), 
         date_final = as.Date(dmy_date, format = "%d-%b-%Y")
         )%>% #remove "temp_" at beginning of each year
  select(-meas_date, -dmy_date, -Date)%>%
  filter(!is.na(date_final)) #remove extra rows with no date

# Wind Cave: 

wind_cave_pivot <- wind_cave %>%
  select(Date, Year, Temp)%>%
  rename(year = Year, windcave_temp = Temp)%>%
  mutate(dmy_date = paste(Date, year, sep = "-"), 
         date_final = as.Date(dmy_date, format = "%d-%b-%Y"), 
         )%>%
  select(-dmy_date, -Date)%>%
  arrange(date_final)

```


Combine into master dataframe: 

```{r}
teton_temps <- merge(middle_teton_pivot, ak_basin_pivot)%>% 
  merge(n_teton_pivot)%>%
  merge(paintbrush_pivot)%>%
  merge(s_cascade_pivot)%>%
  merge(wind_cave_pivot, all.x = T) # Merge all 6 sites; all.x command for final merge keeps all 752 obs for each year. 
```

Calculate Mean summer (July-Aug, based on "real time warming" paper) temperature for each year by site; reformat for easier plotting: 

```{r}
mean_temps <- teton_temps %>%
  mutate(month = month(date_final))%>% #creating month column
  filter(month == 7|month == 8)%>% #filtering for ONLY July and August 
  group_by(year)%>% #group by year
  summarise(
    mid_teton_xbar = mean(mid_teton_temp, na.rm = T), 
    ak_basin_xbar = mean(ak_basin_temp, na.rm = T), 
    n_teton_xbar = mean(n_teton_temp, na.rm = T), 
    paintbrush_xbar = mean(paintbrush_temp, na.rm = T), 
    s_cascade_xbar = mean(s_cascade_temp, na.rm = T), 
    windcave_xbar = mean(windcave_temp, na.rm = T) #calculate average summer temp for each year, removing missing obs.
  )%>%
  pivot_longer(!year, names_to = "site", values_to = "mean_temp")%>% #pivot longer to get one column for site, one for mean temp, one row for each year
  mutate(year = as.numeric(year)) #recode year column as number instead of character
```

Initial data exploration - plotting: 

```{r, message=F, warning=F}
temp.line <- ggplot(mean_temps, aes(x = year, y = mean_temp, color = site))+
  geom_point()+
  geom_line()+
  geom_smooth(se = F, method = "lm", linetype = "dashed", size = 0.5)+
  theme_classic()+
  labs(x = "Year", y = "Mean summer temperature, C", color = "Site Name")+
  scale_color_manual(values = brewer.pal(6, "Dark2"), labels = c("AK Basin", "Middle Teton", "N Teton", "Paintbrush", "S Cascade", "Wind Cave"))
temp.line
```

### B. Testing a linear model for the effects of year within each site: 

First, nest data based on site: 

```{r}
temps_nested <- mean_temps %>%
  nest(data = -site) #create new nested dataframe with one column for site, one column for data - data entries for each site include year and mean temp. 
```

Next, run a model for each site separately and return a summary table with the results for each site:

```{r}
temp_models <- temps_nested %>%
  mutate(model = map(data, ~lm(mean_temp~year, data = .)%>% #run a model of temp~year for each site (referenced by "."); store results in new column called "model"
                     tidy))%>% #tidy function to construct a summary table with model results (seperate tables for each site)
  unnest(model) %>% #unnest model column to get rows for each site, columns for model metrics
  filter(term == "year")  #remove intercept term to just get info on year for each site
temp_models
```

Results: No siginificant trends in any site; non-significant (-) trends in AK basin, N Teton, Paintbrush, and S Cascade; non-significant positive trends in Mid Teton and Wind Cave - makes sense with the model visualization: 

```{r, message = F, warning=F, echo=F}
temp.line
```

### C. Testing a linear model that includes year, snow pack metrics, and air temperature metrics

#### i. Data import and re-formatting:

First, read in the data: Data from the Phillips Bench and Grand Targhee SNOTEL sites is in "teton_snotel.csv" in the teton_trends project directory (downloaded from the NRCS report generator and edited to remove headers, rename columns in Excel for easier import). This file has: 

  * Snow-water equivalent (cm)
  * Snow depth (cm)
  * Air temperature (C)
  
For April 1st 2015 - March 1st 2023 for both sites. On average, peak snowpack for the Tetons is around April 1st (based on graphic from https://www.tetonconservation.org/snowpack-information ), so will try using April 1st SWE and snow depth as predictor variables. 

Data import and some re-formatting: 

```{r}
snotel <- read.csv("teton_snotel.csv")%>% #read in file
  mutate(date = as.Date(date, format = "%m/%d/%y"), #re-code date column as R date obj
         airtemp_c = as.numeric(airtemp_c), #recode air temperature as numeric
         snowdepth_cm = as.numeric(snowdepth_cm)) 
```

Next, calculate average of both sites by grouping by date, calculating overall means: 

```{r}
snotel_xbars <- snotel %>%
  group_by(date)%>% # group by date
  summarise(swe_xbar  = mean(swe_cm, na.rm = T), 
            snowdepth_xbar = mean(snowdepth_cm, na.rm = T),
            airtemp_xbar = mean(airtemp_c, na.rm = T))%>% #calculate means across both sites for each date
 mutate(year = year(date), 
        month = month(date), 
        day = day(date)) #add columns with year, month, and day
```

End goal = data frame with year, mean summer (July-Aug) air temp, April 1st SWE and April 1st snow depth. Two steps: 

  a. Get April 1st SWE and snow depth for each year: 

```{r}
snow <- snotel_xbars %>%
  filter(month == 4 & day == 1)%>% #filter out observations for april 1st of each year
  select(year, swe_xbar, snowdepth_xbar)%>% #select year, swe, and snow depth measurements; drop other cols
  rename(swe_apr1 = swe_xbar, snowdepth_apr1 = snowdepth_xbar) #rename swe and snowdepth cols 
```

  b. Calculate mean summer air temp for each year: 
  
```{r}
air <- snotel_xbars %>%
  filter(month == 7 | month == 8) %>% #filter for just July and August 
  group_by(year) %>% #group by year
  summarise(summertemp_xbar = mean(airtemp_xbar))%>% #calculate mean summer air temp
  select(year, summertemp_xbar) #select year and summer temp for easier merge
```

  c. Merge these to get final snotel data, then merge that dataframe with the mean summer stream temps calculated above: 
  
```{r}
snotel_final <- merge(snow, air) # merge snowpack and air temp info

summer_complete <- merge(mean_temps, snotel_final) #add snowpack and temp info to mean_temps dataframe
```

#### ii. On to modeling: 

First pass: what are the visual relationships between stream temperature and SWE, Snow Depth, and summer air temp? What are the trends in these variables over time? 

```{r, message = F, warning=F}
stream.swe <- ggplot(summer_complete, aes(x = swe_apr1, y = mean_temp, color = site))+
  geom_point()+
  geom_smooth(se = F, method = "lm")+
  theme_classic()+
  labs(x = "April 1st SWE, cm", y = "Mean summer stream temperature, C", color = "Site Name")+
  scale_color_manual(values = brewer.pal(6, "Dark2"), labels = c("AK Basin", "Middle Teton", "N Teton", "Paintbrush", "S Cascade", "Wind Cave"))
stream.swe

stream.sd <- ggplot(summer_complete, aes(x = snowdepth_apr1, y = mean_temp, color = site))+
  geom_point()+
  geom_smooth(se = F, method = "lm")+
  theme_classic()+
  labs(x = "April 1st Snow Depth, cm", y = "Mean summer stream temperature, C", color = "Site Name")+
  scale_color_manual(values = brewer.pal(6, "Dark2"), labels = c("AK Basin", "Middle Teton", "N Teton", "Paintbrush", "S Cascade", "Wind Cave"))
stream.sd

stream.air <- ggplot(summer_complete, aes(x = summertemp_xbar, y = mean_temp, color = site))+
  geom_point()+
  geom_smooth(se = F, method = "lm")+
  theme_classic()+
  labs(x = "Mean July-Aug. Air temp, C", y = "Mean summer stream temperature, C", color = "Site Name")+
  scale_color_manual(values = brewer.pal(6, "Dark2"), labels = c("AK Basin", "Middle Teton", "N Teton", "Paintbrush", "S Cascade", "Wind Cave"))
stream.air
```

First pass: looks like there may be a fairly strong (-) correlation between April 1st SWE/Snow depth and stream temp at N teton and Paintbrush, but probably not in other sites.

Modeling using same approach as above: 

Checking for collinearity in predictor variables: 

```{r}
covariates <- summer_complete %>%
  select(summertemp_xbar, swe_apr1, snowdepth_apr1) # New dataframe with just covariates of interest

M <- cor(covariates) #calculate correlation matrix with the "cor" function (stats package)

corrplot(M, method = "number") #plot the correlation matrix with the corrplot package
```

Strong correlation between snow depth and swe (no surprise), so can only include one at a time. 


**Modeling**:

First, nest data based on site: 

```{r}
summer_nested <- summer_complete %>%
  nest(data = -site) #create new nested dataframe with one column for site, one column for data - data entries for each site include year and mean temp. 
```

Next, run a model for each site separately and return a summary table with the results for each site:

```{r}
summer_models <- summer_nested %>%
  mutate(model = map(data, ~lm(mean_temp~swe_apr1+summertemp_xbar, data = .)%>% #run a model of temp~swe and summer temp for each site (referenced by "."); store results in new column called "model"
                     tidy))%>% #tidy function to construct a summary table with model results (seperate tables for each site)
  unnest(model) %>% #unnest model column to get rows for each site, columns for model metrics
  filter(term == "swe_apr1"|term == "summertemp_xbar")  #remove intercept term to just get info on swe and summer temp for each site
summer_models
```

Marginally significant negative correlation between APR 1st SWE and mean summer temperature at Paintbrush (p = 0.09), not at any other sites; no significant correlations between summer air temp and summer stream temp. 

#### iii. Tangent - any trends in snow depth, swe, or summer air temp between 2015-2022? 

Plots: 

```{r, message = F, warning=F}
swe.time <- ggplot(snotel_final, aes(x = year, y = swe_apr1))+
  geom_point()+
  geom_smooth(se = F, method = "lm")+
  theme_classic()+
  labs(x = "Year", y = "April 1st SWE, cm")
swe.time

sd.time <- ggplot(snotel_final, aes(x = year, y = snowdepth_apr1))+
  geom_point()+
  geom_smooth(se = F, method = "lm")+
  theme_classic()+
  labs(x = "Year", y = "April 1st Snow Depth, cm")
sd.time

air.time <- ggplot(snotel_final, aes(x = year , y = summertemp_xbar))+
  geom_point()+
  geom_smooth(se = F, method = "lm")+
  theme_classic()+
  labs(x = "Year", y = "Mean July-Aug. Air temp, C")
air.time
```

Looks like there could be a significant increase in air temp, weak decreases in snowpack, but nothing obvious. 

Models: 

**a. SWE:** 

```{r}
swe.lm <- lm(swe_apr1 ~ year, data = snotel_final)
summary(swe.lm)
```

No trend in April 1st SWE

**b. Snow depth:** 

```{r}
sd.lm <- lm(snowdepth_apr1 ~ year, data = snotel_final)
summary(sd.lm)
```

No trend in snow depth. 

**c. Summer air temp:** 

```{r}
air.lm <- lm(summertemp_xbar ~ year, data = snotel_final)
summary(air.lm)
```

Significant (+) trend in air temp between 2015-2022 - warming of +0.26 degrees/year. Interesting that increased summer air temp didn't result in increased summer stream temp... 


##2. Comparisons with invert data: 

### A. Testing lm's of richness ~ stream temp

Bring in richness data:

```{r}
richness <- read.csv("invert_data/cleaned_csv/invert_richness.csv")%>%
  rename("site" = "Stream", "year" = "Year")%>%
  select(-X)
```

Add to avg. summer stream temp data: 

```{r}
temps_rename <- mean_temps %>%
  mutate(site = str_remove(site, "_xbar"))%>%
  mutate(site = ifelse(site == "ak_basin", "s_ak_basin", site))
```

Merge: 

```{r}
temps_richness <- merge(richness, temps_rename, all = T)%>%
  filter(!(site == "paintbrush"|site == "s_teton"), !is.na(richness))
```

##3. Re-doing initial analyses with full datasets: 

Read in data:

```{r}
daily_temps <- read.csv("Temperature/cleaned_full_datasets/temps_daily.csv")
```

### A. TASR Sites only: 

Subset TASR data; add source info:

```{r}
tasr.sites <- c("cloudveil", "delta", "grizzly", "gusher", "mid_teton", "n_teton", "paintbrush", "s_ak_basin", "s_cascade", "s_teton", "skillet", "windcave") #make list of TASR sites

source <- read.csv("source_info.csv")%>%
  rename(site = stream) #read in source info, rename for merge

tasr_temps <- merge(daily_temps, source)%>% #add source info to temp dataset
  filter(site%in%tasr.sites) #extract all data for TASR sites
```

Calculate mean summer (July-Aug) temperature for each site + year: 

```{r}
tasr_means <- tasr_temps %>%
  mutate(mo = month(date1), 
         yr = year(date1))%>%
  filter(mo == 7|mo == 8)%>%
  group_by(site, yr, source, full_name)%>%
  summarise(summer_t = mean(temp_xbar, na.rm = T))
```

Initial plotting: 

```{r}
pres.pal <- c("#C26ED6", "#F89225", "#76D96F")

tasr.summer <- ggplot(tasr_means, aes(x = yr, y = summer_t, color = source))+
  geom_point()+
  geom_line()+
  geom_smooth(se = F, method = "lm", linetype = "dashed", size = 0.5)+
  theme_classic()+
  facet_wrap(~full_name, scales = "free_x")+
  scale_color_manual(values = pres.pal, labels = c("Glacier", "Snowmelt", "Icy Seep"))+
  labs(x = "Date", y = "Mean Summer (July-Aug.) Stream Temperature, C", color = "Stream Source")+
  theme(legend.position = "bottom", 
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), 
        axis.title = element_text(size = 11, face = "bold"), 
        axis.text = element_text(size = 8), 
        legend.text = element_text(size = 10), 
        legend.title = element_text(size = 11, face = "bold"), 
        strip.text = element_text(size = 10, face = "bold"))
tasr.summer
```

Save:

```{r}
ggsave("pres_figures/tasr_summer.pdf", tasr.summer, width = 12, height = 5, units = "in", device = "pdf", dpi = "retina")
```


#### Modeling summer temp ~ year: 


First, nest data based on site: 

```{r}
tasr_nested <- tasr_means %>%
  nest(data = -site) #create new nested dataframe with one column for Stream, one column for data - data entries for each site include year and richness. 
```

Next, run a model of richness ~ mean temp for each site separately and return a summary table with the results for each site:

```{r}
tasr_models <- tasr_nested %>%
  mutate(model = map(data, ~lm(summer_t~yr, data = .)%>% #run a model of richness~year for each site (referenced by "."); store results in new column called "model"
                     tidy))%>% #tidy function to construct a summary table with model results (seperate tables for each site)
  unnest(model) %>% #unnest model column to get rows for each site, columns for model metrics
  filter(term == "yr")  #remove intercept term to just get info on year for each site
tasr_models
```

No significant trends in summer temperature at any site. 

### B. Air temperature analysis + plotting:

lm:

```{r}
air.lm <- lm(summertemp_xbar ~ year, data = snotel_final)
summary(air.lm)
```




```{r}
airtemp <- ggplot(snotel_final, aes(x = year , y = summertemp_xbar, color = summertemp_xbar))+
  geom_point()+
  geom_line()+
  geom_smooth(se = F, method = "lm", linetype = "dashed", color = "DarkGrey", size = 0.5)+
  theme_classic()+
  labs(x = "Year", y = "Mean Summer (July-Aug.) Air temp, C")+
  scale_color_gradient(low = "#D5A5A3FF", high = "#753835FF")+
  theme(legend.position = "none", 
        axis.title = element_text(size = 11, face = "bold"), 
        axis.text = element_text(size = 10)
        )+
  annotate(geom="label", x = 2015.8, y = 15.75, label = "P = 0.03; R Sq. = 0.49", fontface = "bold", size = 4)
airtemp
```

Save:

```{r}
ggsave("pres_figures/air_summer.pdf", airtemp, width = 12, height = 5, units = "in", device = "pdf", dpi = "retina")
```

