---
title: "August model results"
author: "Justin Pomeranz"
date: "2025-05-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(brms)
library(tidybayes)

fit_data_orig <- readRDS(
  here::here("Temperature/brms_models/all_sites/august/hurdle_fit_data.rds"))
brm1 <- readRDS(here::here("Temperature/brms_models/all_sites/august/fit_rand_slopes_hurdle_aug_rand_site.rds"))

fit_data_year_summary <- readRDS(
  here::here("Temperature/brms_models/all_sites/August/fit_data_year_summary.rds"))

# rmarkdown::render("Temperature/brms_models/all_sites/august/Aug_results_2025-05.Rmd")
```

## Overview  

This document summarizes the model fitting and posterior for mean stream temperatures in August for the Teton Alpine Stream Research (TASR) project. 


## Setup  

* Load libraries  

```{r, eval=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(brms)
library(tidybayes)

```

* Load data  
* Data has already been modified in the following ways:  
  * Only data for the month of August is included  
  * `year_s` has been centered at 0 and scaled by the `sd(year)`  
  * `temp_1` has been scaled to be from 0-1 as: $ \text{temp}_1 = \text{temp}_i / max(\text{temp})$  
  
  
```{r, eval=FALSE}
fit_data_orig <- readRDS(
  here::here("Temperature/brms_models/all_sites/august/fit_data.rds"))
fit_data_year_summary <- readRDS(
  here::here("Temperature/brms_models/all_sites/August/fit_data_year_summary.rds"))
```

* Load model  

```{r, eval=FALSE}
brm1 <- readRDS(here::here("Temperature/brms_models/all_sites/august/fit_rand_slopes_hurdle_aug_rand_site.rds"))
```

## Model details  

* 4 chains with 2000 iterations  
  * half discarded as warmup  
  * 4000 posterior iterations  
  
* The model has the following formula  

```{r}
brm1$formula
```

Where `source` is a categorical factor with the following levels: `rock_glacier`, `glacier`, and `snowfield`.  



## Posterior Prediction  

* Density overlay for sources  

```{r, warning=FALSE, message=FALSE}
pp_check(brm1, 
         type = "dens_overlay_grouped",
         group = "source") +
  scale_x_log10() +
  theme_bw()
```

* boxplot of posterior  
* Note that this model correctly captures the lower bound of 0, and does not predict negative numbers.  

```{r}
pp_check(brm1, 
         type = "boxplot") +
  theme_bw()
```

## Bayesian R^2  

```{r}
bayes_R2(brm1)
```

* This model explains 63.6% (CrI: 63.1, 64.1) of the variation in August temperatures.  

## Model Posteriors  

* Now we will explore the model posteriors to understand the effects of stream source on August temperatures  

### Source effect  

#### Data for prediction  

* First, calculate the average and SD of year from the original data, and the maximum temperature  
* These are used to back calculate model posteriors to original scales as appropriate.  
```{r}
mean_year <- fit_data_year_summary$mean_year
sd_year <- fit_data_year_summary$sd_year
max_temp <-  max(fit_data_orig$temp_c)
```


* Make a data frame with the original:  
  * `source`  
  * `year_s` (Z_score of year)  

* We will then make a date of August 15th for each year for plotting purposes  

* This data will be used to add posterior draws based on the model in the next step  

```{r}
source_data <- brm1$data |>
  select(source, year_s) |> 
  distinct() |> # all combinations of source and year_s from empirical data
  mutate(year = (year_s*sd_year)+mean_year, # back calculate year Z-score to original year  
         month = 8, # manually set month number
         day = 15, # and day number
         date = make_date(year, month, day)) |> # set this as a `date` variable type for plotting
  select(source, year_s, date) # only keep variables we need later

head(source_data) # see first few rows as an example
```
* Visualize the effect of year and source on August stream temperatures  

```{r}
# add posterior draws 
add_epred_draws(
  newdata = source_data, # new data made above to add expected (mean) temperature values based on the posterior
  brm1, # the model fit
  re_formula = NA) |> # ignores random effects of site and just shows the overall fixed main and interactive effects of site*year
  mutate(.epred = (.epred*max_temp)) |> # converts back to original temperature scale  
  ggplot(aes(x = date, 
             y = .epred,
             fill = source,
             color = source)) +
  stat_lineribbon(aes(y = .epred),
                  .width = c(0.95, 0.68),
                  alpha = 1/4) +
  theme_bw() +
  labs(title = "Mean August Temperatures by Source",
       subtitle = "Ribbons = 68 and 95% CrI",
       x = "Year",
       y = "Temperature") +
  facet_wrap(~source) +
  scale_fill_manual(values = c("deepskyblue",
                               "slategrey", 
                               "springgreen4")) +
  scale_color_manual(values = c("deepskyblue",
                                "slategrey", 
                                "springgreen4"))
```

## Magnitude of change per year  

* Calculate the change from year to year  
* This is a Gamma model, so the interpretation of the "slope" ($\beta_{year}$) coefficient is different from a standard generalized linear model  
  * Instead of saying "Temperatures increase by "x" each year", we say "Temperatures chnage each year by "x percentage".  
  * This is why there is a "curve" in the plots above.  
  
* In order to estimate the percent change, I will create a dataframe with only 2 years from the middle of the data.  
  * The expected values will be added with `add_epred_draws()`  
  * The absolute difference will be calculated by subtracting the value in the last year from the first year  
  * The percent change will then be calculated by dividing the difference by the value in the first year times 100  
  * The distribution of the percent changes will be reported.  

* Using other year-pairs from the data does not change the results (SI).    

```{r}
# 2019 and 2020
y_19_20 <- data.frame(
  year_real = c(2019, 2020)) |>
  mutate(year_s = (year_real - mean_year)/sd_year) |>
  expand_grid(source = unique(brm1$data$source)) |>
  add_epred_draws(brm1,
                  re_formula = NA,
                  allow_new_levels = TRUE) |>
  ungroup()|>
  mutate(.epred = .epred * max_temp) |>
  select(-.row, -.chain, -.iteration, -year_s) |>
  pivot_wider(names_from = "year_real",
              values_from = ".epred") |>
  mutate(slope_new = `2020` - `2019`) |>
  # slope_new = chnage in actual temperature right scale
  mutate(prop = slope_new / `2019` * 100)

y_19_20 |>
  group_by(source) |>
  mean_qi(prop)
```

* `snowfield` streams increase by an average of 5.28% (CrI 0.64, 10)  
  * 95% CrI is entirely positive (see halfeye plots below)  

* `glacier` streams increase by an average of 3.51% (CrI -1.6, 9)  
  * 95% CrI is mostly positive, but there are also negative estimates  

* `rock_glacier` streams increase by an average of 2.47% (CrI -1.52, 6.89)    
  * 95% CrI is entirely positive (see halfeye plots below)  
  

## Probability of positive slopes?  

* The mean magnitude of the percent change form year to year is reported above  
* But what is the probability that these slopes are positive?  

* Calculate the proportion of slopes which are positive  
* make a halfeye plot with the positive iterations highlighted  

### Pr(positive)  

```{r}
pr_pos <- y_19_20 |>
  group_by(source) |>
  mutate(pos = slope_new >0) |>
  summarize(pr_pos = mean(pos))

pr_pos
```

* `snowfield` streams have a 98.4% chance of having increasing mean August temperatures  
* `glacier` streams have a 93.6% chance of having increasing mean August temperatures  
* `rock_glacier` streams have an 89.8% chance of having increasing mean August temperatures  

### Halfeye plot  
```{r}
y_19_20 |>
  ggplot(
    aes(y = source,
        x = slope_new,
        fill = after_stat(x >0))) +
  stat_halfeye() +
  geom_vline(xintercept = 0, 
             linetype = "dashed") +
  scale_fill_manual(values = c("gray80", "orangered1")) +
  theme_bw() +
  labs(x = "Posterior distribution of percent change per year") +
  annotate("text", # rock_glacier
           x = 0.5,
           y = 2.5,
           label = paste("P(Positive)",
                         round(pr_pos[2,2],2),
                         sep = " ")) +
  annotate("text", # Glacier
           x = 0.5,
           y = 1.5,
           label = paste("P(Positive)",
                         round(pr_pos[1,2],2),
                         sep = " ")) +
  annotate("text", # snow
           x = 0.5,
           y = 3.5,
           label = paste("P(Positive)",
                         round(pr_pos[3,2],2),
                         sep = " ")) +
  NULL

```

## Site-specific predictions  

* Here, we have the results for each site  

* First, make site-level data  
* add expected posterior draws  
* summarize and plot results  

```{r}
site_data <- brm1$data |>
  select(source, year_s, site) |>
  distinct() |>
  group_by(site) |>
  # filter(year_s == min(year_s) | 
  #          year_s == max(year_s)) |>
  mutate(year = (year_s*sd_year)+mean_year,
         month = 8,
         day = 15,
         date = make_date(year, month, day))

site_temps <- add_epred_draws(
  newdata = site_data,
  brm1,
  #allow_new_levels = TRUE,
  re_formula = NULL) |> # includes random effects
  mutate(.temp = (.epred*max_temp)) 

site_temps
```

### Plot through time  

```{r}
ggplot(site_temps, 
       aes(x = date, 
           y = .epred,
           fill = site,
           color = site)) +
  stat_lineribbon(aes(y = .epred),
                  .width = c(0.95, 0.68),
                  alpha = 1/4) +
  theme_bw() +
  labs(title = "Mean August Temperatures by Source",
       subtitle = "Ribbons = 68 and 95% CrI",
       x = "Year",
       y = "Temperature") +
  facet_wrap(~source) 
```

### Site-specific percent change  

```{r}
s_19_20 <- brm1$data |> 
  select(source, site) |>
  distinct() |>
  expand_grid(year_real = c(2019, 2020)) |>
  mutate(year_s = (year_real - mean_year)/sd_year) |>
  add_epred_draws(brm1,
                  re_formula = NULL) |>
  ungroup()|>
  mutate(.epred = .epred * max_temp) |>
  select(-.row, -.chain, -.iteration, -year_s) |>
  pivot_wider(names_from = "year_real",
              values_from = ".epred") |>
  mutate(slope_new = `2020` - `2019`) |>
  # slope_new = change in actual temperature right scale
  mutate(prop = slope_new / `2019` * 100)

s_19_20 |>
  group_by(site, source) |>
  mean_qi(prop) |>
  arrange(source, site, prop)
```


### Site-specific Pr(+)  

```{r}
s_19_20 |>
  group_by(site) |>
  mutate(pos = slope_new >0) |>
  summarize(pr_pos = mean(pos))
```

### site-specific change in temperatures  

* mean temperature change  
* first year with empirical data  
* last year with empirical data (mostly 2024)  

```{r}
brm1$data |>
  select(source, year_s, site) |>
  distinct() |>
  group_by(site) |>
  filter(year_s == min(year_s) |
           year_s == max(year_s)) |>
  mutate(year = (year_s*sd_year)+mean_year) |>
  add_epred_draws(
  brm1,
  re_formula = NULL) |> # includes random effects
  mutate(.temp = (.epred*max_temp)) |>
  ungroup() |>
  select(site, year, .temp, .draw, source) |>
  group_by(source, site, .draw) |>
  mutate(year_cat = case_when(
    year == min(year) ~ "start", 
         .default = "end")) |>
  select(-year) |>
  pivot_wider(id_cols = c(site, .draw, source),
              names_from = year_cat,
              values_from = .temp) |>
  mutate(delta = end - start,
         group = cur_group_id()) |>
  ggplot(aes(y = fct_reorder(site, group), 
             x = delta, 
             fill = source)) +
  stat_halfeye()+
  scale_fill_manual(values = c("deepskyblue",
                               "slategrey", 
                               "springgreen4")) +
  theme_bw() +
  geom_vline(xintercept = 0, 
             linetype = "dashed") +
  labs(y = "site",
       x = "Absolute change in average August temperatures")
```

* range of years for each site  

```{r}
site_temps |>
  ungroup() |>
  select(site, year) |>
  group_by(site) |>
  filter(year == min(year) |
           year == max(year)) |>
  distinct(site, year) |>
  mutate(year_cat = case_when(
    year == min(year) ~ "start", 
         .default = "end")) |>
  pivot_wider(id_cols = c(site),
              names_from = year_cat,
              values_from = year) |>
  mutate(year_range = end - start) |>
  arrange(start, end)
```


# SI  

## Model Fit  

### Model MCMC Chains and posterior of parameters  

```{r}
plot(brm1, ask = FALSE)
```

### All Rhats < 1.1  

```{r}
summary(brm1)
```

## Percent change  

### Comparison between other year pairs  

#### 2015 to 2016  
```{r}
data.frame(
  year_real = c(2015, 2016)) |>
  mutate(year_s = (year_real - mean_year)/sd_year) |>
  expand_grid(source = unique(brm1$data$source)) |>
  add_epred_draws(brm1,
                  re_formula = NA,
                  allow_new_levels = TRUE) |>
  ungroup()|>
  mutate(.epred = .epred * max_temp) |>
  select(-.row, -.chain, -.iteration, -year_s) |>
  pivot_wider(names_from = "year_real",
              values_from = ".epred") |>
  mutate(slope_new = `2016` - `2015`) |>
  # slope_new = chnage in actual temperature right scale
  mutate(prop = slope_new / `2015` * 100) |>
  group_by(source) |>
  mean_qi(prop)

```



#### 2023 to 2024  

```{r}
data.frame(
  year_real = c(2023, 2024)) |>
  mutate(year_s = (year_real - mean_year)/sd_year) |>
  expand_grid(source = unique(brm1$data$source)) |>
  add_epred_draws(brm1,
                  re_formula = NA,
                  allow_new_levels = TRUE) |>
  ungroup()|>
  mutate(.epred = .epred * max_temp) |>
  select(-.row, -.chain, -.iteration, -year_s) |>
  pivot_wider(names_from = "year_real",
              values_from = ".epred") |>
  mutate(slope_new = `2024` - `2023`) |>
  # slope_new = chnage in actual temperature right scale
  mutate(prop = slope_new / `2023` * 100) |>
  group_by(source) |>
  mean_qi(prop)

```

* Exact same results as 2019-2020 reported above  
